|

### 2024-09-05 manuscript/paperCMSACversion.Rmd
more editing

We note than in @Popkin2023, they find 183 E-P enforcers; we have only 182 in our data set with the discrepancy due the the removal of the 4 players with missing height information.

@@

These results suggest that the conclusions of @Popkin2023 are overstated due to the differences in exposure time between the groups at the time of their final analysis.

join all from jama nhl data, condition on death from jama deaths and look at yob between enforcer/control
```{r}
cf = read.csv('../JAMA NHL Data/C-F-Table 1.csv')
cp = read.csv('../JAMA NHL Data/C-P-Table 1.csv')

@@

jama_short %>%
  filter(Player %in% jama_deaths$Enforcers) %>%
  mutate(enforcer = ifelse(EF == 1 | EP == 1, 1, 0)) %>% 
  group_by(enforcer) %>%
  summarize(mean(DOB.Year))

jama_short %>%
  mutate(deathage = as.numeric(difftime(Death.Date,
                                        Birth.Date,
                                        units = "weeks")/52.25)) %>%
  filter(Player %in% jama_deaths$Enforcers) %>%
  mutate(enforcer = ifelse(EF == 1 | EP == 1, 1, 0)) %>% 
  group_by(enforcer) %>%
  summarize(mean(deathage))

@@

While we do not address this imbalance here, as we focus on closely following the work of @Popkin23, it is worth noting the substantial imbalance in geographic origin (i.e. North American vs outside of North America) of enforcers versus non-enforcer.

>>>>

We note than in @Popkin2023, they find 183 E-P enforcers; we have only 182 in our data set with the discrepancy due the the removal of the 4 players with missing height information, which is necessary to perform the survival analysis.

@@

```{r popkin_conclusions, echo = FALSE, tab.cap = ""}
cf = read.csv('../JAMA NHL Data/C-F-Table 1.csv')
cp = read.csv('../JAMA NHL Data/C-P-Table 1.csv')

@@

cnclsn_sum = jama_short %>%
  filter(Player %in% jama_deaths$Enforcers) %>%
  mutate(enforcer = ifelse(EF == 1 | EP == 1, 1, 0), 
         deathage = as.numeric(difftime(Death.Date,
                                        Birth.Date,
                                        units = "weeks")/52.25)) %>% 
  group_by(enforcer) %>%
  summarize(mean_YOB = mean(DOB.Year), mean_age_at_death = mean(deathage))

```
<!-- latex table generated in R 4.4.1 by xtable 1.8-4 package
Wed Sep  4 18:44:57 2024-->

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & enforcer & mean\_YOB & mean\_age\_at\_death \\ 
  \hline
1 & 0.00 & 1958.29 & 56.46 \\ 
  2 & 1.00 & 1962.81 & 47.15 \\ 
   \hline
\end{tabular}
\caption{Original Paper Conclusions Data Summary
 \label{tab:popkin_conclusions}}
\end{table}

These results suggest that the conclusions of @Popkin2023 are overstated due to the differences in exposure time between the groups at the time of their final analysis. Table \ref{tab:popkin_conclusions}

@@

While we do not address this imbalance here, as we focus on closely following the work of @Popkin2023, it is worth noting the substantial imbalance in geographic origin (i.e. North American vs outside of North America) of enforcers versus non-enforcer.
